\vspace{-.2in} Squint: Lossy Hierarchical Compression on Symbol Streams \vspace{-.1in}

### 2012-12-12 tex/paper.tex
typos

It may seem strange at first to allow a string such as "mind" to be lossily
corrupted into, say, ``rind''.  If, however, one is more interested in
recurrent structure in the text than in its precise semantic meaning, these
corruptions are unimportant.  For instance, when looking at text in a foreign
language, one is more likely to pick up the texture of the passage, the
repeated sounds, common phrases, and breaks between larger-scale sections of
text.  In this situation, a change of a single letter would probably go
unnoticed.  Similarly, when skimming through a huge file, say, an execution
log, an administrator usually begins by looking for patterns that stand out.
Lorem ipsum is a perfect example of a piece of ``text'' whose value lies in its
texture and sentence/paragraph structure, and we will return to it in later
sections.

@@

Kieffer briefly identifies two primary methods for lossy hierarchical
compression: wavelet-based and fractal-based schemes \cite{tutorial}.  In
wavelet-based schemas, a signal is compressed by recording the largest
coefficients of its wavelet decomposition, in effect trying to represent a
signal as closely as possible using wavelet building blocks.  The JPEG image
format uses a similar technique in order to compress images lossily
\cite{jpeg}.  Fractal based compression takes this a step further and
represents parts of a signal using self-similarity, encoding an image for
instance as a fixed point of contraction maps.  This is similar in spirit to
what is done in MPEG-1, where parts of one frame can be transformed to encode
parts of the next frame. Many of these ideas, such as finding largest matching
components of objects, make their way into our algorithms.

@@

\begin{algorithm}[h]
\caption{Similarity Lossifier Algorithm}
\label{sim_alg}
\begin{algorithmic}[1]
\Procedure{Lossify-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \For{$var_1 \rightarrow rhs_1,\ var_2 \rightarrow rhs_2 \in g$}
      \State $str_{i} \gets expand(rhs_i)$
      \If{$lsim(str_{1},str_{2})<\epsilon$}
        \State replace the larger var with the smaller throughout $g$
        \State break
      \EndIf
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

@@

The previous algorithm updates the grammar sequentially, since each variable
replacement can have an effect on large parts of the grammar, we will have to
go through comparing all pairs again after the change.  An alternative
algorithm would be to try to replace as many similar variables as possible
all in one sweep. This would
avoid many of the repeat pair-comparisons done in \Similarity. In other
words, the algorithm below puts similar variables into equivalence class
clusters, and then replaces each variable in a cluster them with the smallest
representative all at once each iteration.

\begin{algorithm}[h]
\caption{Cluster Lossifier Algorithm}
\label{cluster_alg}
\begin{algorithmic}[1]
\Procedure{Cluster-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \State $clusters \gets ()$
    \For{$var_1 \rightarrow rhs_1 \in g$}
      \For{$c \in clusters$}
        \State $(var_2 \rightarrow rhs_2) \gets $ first variable in $c$
        \State $str_{i} \gets expand(rhs_i)$
        \If{$lsim(str_{1},str_{2})<\epsilon$}
          \State add $var_1$ to $c$
          \State break
        \EndIf
      \EndFor
      \If{$var_1$ hasn't been added to a cluster}
        \State add $var_1$ to a new cluster in $clusters$
      \EndIf
    \EndFor
    \State save a mapping $m$ from variables to sizes
    \For{Cluster $c \in clusters$}
      \State replace all $var \in c$ with the smallest (smallest w.r.t. $m$).
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

Both of the algorithms depend on the concept of the "smallest / smaller"
variable. We can define the size of a variable to be the length of the
expansion of its rhs. In this way, if we think of our CFG's as a DAG, then
variables higher up on the DAG will always have larger expansions than those
lower down, and thus will have a larger size. By always replacing with the
smallest variable, we avoid creating cyclic references in our CFG.  We will
prove this fact in the next section.

@@

We can view the CFG as a DAG where for variables $V_1$, $V_2$, we add an edge
whenever $V_1$ contains $V_2$ in its rhs. When sizes are computed on a CFG
before we modify it, if $V_1 \rightarrow V_2$ then $size(V_1) > size(V_2)$
since the rhs of $V_1$ contains $V_2$.  Replacing a variable $V_1$ with $V_2$
has the same effect as redirecting all in-edges to $V_1$ onto $V_2$.

@@

For the \Cluster algorithm the analysis is a bit more involved, since to avoid
recomputation we only calculate the sizes of the variables once before doing
all of the replacements. First note that even as we modify the DAG, all of the
edges we create in the DAG will respect the ordering of sizes that we
calculated before we did any replacements. This is because each time we
redirect an edge from $X\rightarrow A$ to $X\rightarrow B$
$size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.  Suppose
then that we created a cycle during \Cluster by redirecting an edge from $X
\rightarrow A$ to $X \rightarrow B$.  This means that there must have been a
path from $B$ to $X$. However, even in its intermediate state we know that the
DAG respects the $size$ ordering, and $size(X) > size(A) > size(B)$, so there
cannot be a path from $B$ to $X$. This is a contradiction and we are done.

@@

From the \Sequitur paper \cite{sequitur} we know that the number of rules in
the grammar generated by \Sequitur is bounded by $O(n)$. Thus the worst
case runtime on the similarity algorithm is $\boxed{O(n^4)}$. This is because
we could have up to $r$ replacements.  However, in practice we will see that
the runtime is much better than $O(n^4)$. There are a variety of possible
reasons for this: in our test data, $r$ didn't grow quite linearly in $n$, $R$
could grow much slower than $r$, and the use of native code in calculating the
Levenshtein distance made the its contribution much less noticeable for the
input sizes we tested.

@@

The lossifier algorithm reduced the string to a pair of pairs of the same
equations, where each equation consists of the sum of triples of "12" or "32".
Smaller discrepencies in the numbers involved were dropped, while the filler
commas and linebreaks also ended up being lost. Note that the structure
described is easily observable from the grammar generated.

@@

In table \ref{simclus_r_t} we report the results on running \Sequitur and
\Cluster, without the grammar reductions, on the rep-n datasets. In table
\ref{simclus_g_t} we do the same for the gen-n datasets.  In these tests we
were looking to evaluate the relative fidelity, compression ratios, and runtime
of the two algorithms.

In the tables, Time is the wall clock running as measured on a 2.26 Ghz Intel
Core 2 Duo Macbook Pro.  Ratio is the relative size of the grammar produced
compared with \Sequitur, where the size of the grammar is the sum of the
lengths of the rhs's of the rules. Levdistance is the Levenshtein distance
between the output of the lossified grammar and the original input string, it
is a kind of measure of the fidelity of our algorithms but doesn't take into
account how well patterns were preserved. For the rep-n tests, proportion is
the proportion of the final string produced that was `*'. The closer to
1/30~0.033, the more accurate the lossifier was in recreating the original
texture of the input.

@@

In both datasets, the compression ratios achieved for both algorithms are
roughly equal given the same value for $\epsilon$, so it is fair to compare the
algorithms for the same values of $\epsilon$.  From plotting the data, we see
that runtime of the algorithms are roughly polynomial in the size of the data,
so we can perform a linear regression on the log-log transform of the data to
find the exponents $\alpha$ for the runtime $r(n)=n^\alpha$ where $n$ is the
length of the input string. In the charts below, all of the linear regressions
had $R^2$ values of over $0.98$.
\begin{align*}
  \text{rep-n:}\quad
  \begin{array}{l|l|l||l|l}
    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
    \epsilon & .4 & .2 & .4 & .2 \\
    \alpha   & 1.859 & 2.026 & 1.306 & 1.305
  \end{array}
  &&
  \text{gen-n:}\quad
  \begin{array}{l|l|l||l|l}
    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
    \epsilon & .4 & .2 & .4 & .2 \\
    \alpha   & 2.629 & 2.493 & 1.814 & 2.061
  \end{array}
\end{align*}

@@

The actual compression ratio was respectable in most tests, with both
\Similarity and \Cluster coming in at nearly half the size of the \Sequitur CFG
for the rep-n data, and 70 percent for the gen-n data. A notable exception was
the gen-n test with $\epsilon=.2$, where $\epsilon$ was so small that barely
any substitutions were possible.  However, setting higher values for $\epsilon$
had diminishing returns for compression for larger file sizes, at the cost of
drastically lowering the fidelity of the output.  Data with differing amuounts
of repetition and noise appear to require different values of $\epsilon$ for
the optimal fidelity/compression ratio.

@@

Although the grammars generated by \Sequitur are irreducible, our lossifiers
produce grammars which could potentially be compressed further (losslessly)
using Reduction Rules without changing the output string. To test these rules,
we looked at the rep-n and gen-n test cases again, this time focusing on the
clustering algorithm for $\epsilon = .2$. We report our results in table
\ref{red_t}, where we give the ratio on top of the simplified grammar produced
by the clustering algorithm, and the time taken to run just the reduction
rules.

When plotted on a log-log scale, the runtime for our straightforward
implementation of the reduction rules appears to be superpolynomial. While
modest additional gains of about 30 percent are possible for highly
self-similar data, very few additional gains are possible for more diverse
datasets such as gen-n. This appears to be because the only simplifications
that are possible on diverse datasets don't lead to the wide-ranging repetitive
substitutions which allow reduction rules to be allowed.

@@

The second piece of text is one of the entries to this year's International
Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c}
. We believe this is a particularly good example since its structure is
foreign, but still discernable. Whitespace is particularly important in
standard english but our lossifiers don't treat it any differently from any
other symbol.  The size of the lossified grammar is 1576 compared with the
original size of 1973. The intricate whitespacing was lost, but the clear
delineation into sections was kept, key words such as `define' and `for' are
very visible, and the texture of the obfuscated C is very obvious. For
instance, after the distinct `\#define' headers, the first sections has a bunch
of control statements like `for' and `if', while the middle sections have very
generous spacing and many repeated pointer derence operators such as `***'.
Both of these kinds of textures are retained after lossifying.

@@

1:9 And God said, Let the waters under the heaven be gathered together
unto one place, and let the dry land appear: and it was so.

1:10 And God called the dry land Earth; and the gathering together of
the waters called he Seas: and God saw that it was good.

@@

>>>>

It may seem strange at first to allow a string such as ``mind'' to be lossily
corrupted into, say, ``rind''.  If, however, one is more interested in
recurrent structure in the text than in its precise semantic meaning, these
corruptions are unimportant.  For instance, when looking at text in a foreign
language, one is more likely to pick up the texture of the passage, the
repeated sounds, common phrases, and breaks between larger-scale sections of
text.  In this situation, a change of a single letter would probably go
unnoticed.  Similarly, when skimming through a huge file, say, an execution
log, an administrator usually begins by looking for patterns that stand out.
Lorem ipsum is a perfect example of a piece of ``text'' whose value lies in its
texture and sentence/paragraph structure, and we will return to it in later
sections.

@@

Kieffer briefly identifies two primary methods for lossy hierarchical
compression: wavelet-based and fractal-based schemes \cite{tutorial}.  In
wavelet-based schemes, a signal is compressed by recording the largest
coefficients of its wavelet decomposition, in effect trying to represent a
signal as closely as possible using wavelet building blocks.  The JPEG image
format uses a similar technique in order to compress images lossily
\cite{jpeg}.  Fractal based compression takes this a step further and
represents parts of a signal using self-similarity, encoding an image for
instance as a fixed point of contraction maps.  This is similar in spirit to
what is done in MPEG-1, where parts of one frame can be transformed to encode
parts of the next frame. Many of these ideas, such as finding largest matching
components of objects, make their way into our algorithms.

@@

\begin{algorithm}[h]
\caption{Similarity Lossifier Algorithm}
\label{sim_alg}
\begin{algorithmic}[1]
\Procedure{Lossify-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \For{$var_1 \rightarrow rhs_1,\ var_2 \rightarrow rhs_2 \in g$}
      \State $str_{i} \gets expand(rhs_i)$
      \If{$lsim(str_{1},str_{2})<\epsilon$}
        \State replace the larger var with the smaller throughout $g$
        \State break
      \EndIf
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\label{s_alg}
\end{algorithm}

@@

The previous algorithm updates the grammar sequentially, since each variable
replacement can have an effect on large parts of the grammar, we will have to
go through comparing all pairs again after the change.  An alternative
algorithm would be to try to replace as many similar variables as possible
all in one sweep. This would
avoid many of the repeat pair-comparisons done in \Similarity. In other
words, the \Cluster puts similar variables into equivalence class
clusters, and then replaces each variable in a cluster them with the smallest
representative all at once each iteration. More detailed pseudocode is given
in Algorithm \ref{c_alg} below.

\begin{algorithm}[h]
\caption{Cluster Lossifier Algorithm}
\label{cluster_alg}
\begin{algorithmic}[1]
\Procedure{Cluster-Sim}{$g$} \Comment{Simplify grammar g}
  \While{there are similar variables to replace}
    \State $clusters \gets ()$
    \For{$var_1 \rightarrow rhs_1 \in g$}
      \For{$c \in clusters$}
        \State $(var_2 \rightarrow rhs_2) \gets $ first variable in $c$
        \State $str_{i} \gets expand(rhs_i)$
        \If{$lsim(str_{1},str_{2})<\epsilon$}
          \State add $var_1$ to $c$
          \State break
        \EndIf
      \EndFor
      \If{$var_1$ hasn't been added to a cluster}
        \State add $var_1$ to a new cluster in $clusters$
      \EndIf
    \EndFor
    \State save a mapping $m$ from variables to sizes
    \For{Cluster $c \in clusters$}
      \State replace all $var \in c$ with the smallest (smallest w.r.t. $m$).
    \EndFor
  \EndWhile
\EndProcedure
\end{algorithmic}
\label{c_alg}
\end{algorithm}

Both of the algorithms depend on the concept of the ``smallest / smaller''
variable. We can define the size of a variable to be the length of the
expansion of its rhs. In this way, if we think of our CFG's as a DAG, then
variables higher up on the DAG will always have larger expansions than those
lower down, and thus will have a larger size. By always replacing with the
smallest variable, we avoid creating cyclic references in our CFG.  We will
prove this fact in the next section.

@@

We can view the CFG as a DAG where for variables $V_1$, $V_2$, we add an edge
whenever $V_1$ contains $V_2$ in its rhs. When sizes are computed on a CFG
before we modify it, if $V_1 \hookrightarrow V_2$ is an edge then $size(V_1) > size(V_2)$
since the rhs of $V_1$ contains $V_2$.  Replacing a variable $V_1$ with $V_2$
has the same effect as redirecting all in-edges to $V_1$ onto $V_2$.

@@

For the \Cluster algorithm the analysis is a bit more involved, since to avoid
recomputation we only calculate the sizes of the variables once before doing
all of the replacements. First note that even as we modify the DAG, all of the
edges we create in the DAG will respect the ordering of sizes that we
calculated before we did any replacements. This is because each time we
redirect an edge from $X\hookrightarrow A$ to $X\hookrightarrow B$
$size(X)>size(A)>size(B)$ and transitivity will preserve the ordering.  Suppose
then that we created a cycle during \Cluster by redirecting an edge from $X
\hookrightarrow A$ to $X \hookrightarrow B$.  This means that there must have been a
path from $B$ to $X$. However, even in its intermediate state we know that the
DAG respects the $size$ ordering, and $size(X) > size(A) > size(B)$, so there
cannot be a path from $B$ to $X$. This is a contradiction and we are done.

@@

From the \Sequitur paper \cite{sequitur} we know that the number of rules in
the grammar generated by \Sequitur is bounded by $O(n)$. Thus the worst
case runtime on the similarity algorithm is $\boxed{O(n^4)}$. This is because
we could have up to $r$ replacements.  However, in practice we will see that
the runtime is much better than $O(n^4)$. There are a variety of possible
reasons for this: in our test data, $r$ didn't grow quite linearly in $n$, $R$
could grow much slower than $r$, and the use of native code in calculating the
Levenshtein distance made its contribution much less noticeable for the
input sizes we tested.

@@

The lossifier algorithm reduced the string to a pair of pairs of the same
equations, where each equation consists of the sum of triples of ``12'' or ``32''.
Smaller discrepencies in the numbers involved were dropped, while the filler
commas and linebreaks also ended up being lost. Note that the structure
described is easily observable from the grammar generated.

@@

In table \ref{simclus_r_t} in the appendix we report the results on running \Sequitur and
\Cluster, without the grammar reductions, on the rep-n datasets. In table
\ref{simclus_g_t} in the appendix we do the same for the gen-n datasets.  In these tests we
were looking to evaluate the relative fidelity, compression ratios, and runtime
of the two algorithms.

In the tables, Time is the wall clock running as measured on a 2.26 Ghz Intel
Core 2 Duo Macbook Pro.  Ratio is the relative size of the grammar produced
compared with \Sequitur, where the size of the grammar is the sum of the
lengths of the rhs's of the rules. \emph{Lev.} is the Levenshtein distance
between the output of the lossified grammar and the original input string, it
is a kind of measure of the fidelity of our algorithms but doesn't take into
account how well patterns were preserved. For the rep-n tests, \emph{Prop.} is
the proportion of the final string produced that was `*'. The closer to
$1/30 \approx 0.033$, the more accurate the lossifier was in recreating the original
texture of the input.

@@

In both datasets, the compression ratios achieved for both algorithms are
roughly equal given the same value for $\epsilon$, so it is fair to compare the
algorithms for the same values of $\epsilon$.  From plotting the data, we see
that the runtimes of the algorithms are roughly polynomial in the size of the data,
so we can perform a linear regression on the log-log transform of the data to
find the exponents $\alpha$ for the runtime $r(n)=n^\alpha$ where $n$ is the
length of the input string. In the charts below, all of the linear regressions
had $R^2$ values of over $0.98$.
\begin{align*}
  \text{rep-n:}\quad
  \begin{array}{l|l|l||l|l}
    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
    \epsilon & .4 & .2 & .4 & .2 \\
    \alpha   & 1.859 & 2.026 & 1.306 & 1.305
  \end{array}
  &&
  \text{gen-n:}\quad
  \begin{array}{l|l|l||l|l}
    & \multicolumn{2}{c||}{\Similarity} & \multicolumn{2}{c}{\Cluster} \\ \hline
    \epsilon & .4 & .2 & .4 & .2 \\
    \alpha   & 2.629 & 2.493 & 1.814 & 2.061
  \end{array}
\end{align*}

@@

The actual compression ratio was respectable in most tests, with both
\Similarity and \Cluster coming in at nearly half the size of the \Sequitur CFG
for the rep-n data, and 70 percent for the gen-n data. This is promising because
in \cite{sequitur}, the authors show that \Sequitur on its own achieves better
ratios than most standard compression algorithms.
A notable exception was
the gen-n test with $\epsilon=.2$, where $\epsilon$ was so small that barely
any substitutions were possible.  However, setting higher values for $\epsilon$
had diminishing returns for compression for larger file sizes, at the cost of
drastically lowering the fidelity of the output.  Data with differing amuounts
of repetition and noise appear to require different values of $\epsilon$ for
the optimal fidelity/compression ratio.

@@

Although the grammars generated by \Sequitur are irreducible, our lossifiers
produce grammars which could potentially be compressed further (losslessly)
using Reduction Rules without changing the output string. To test these rules,
we looked at the rep-n and gen-n test cases again, this time focusing on the
clustering algorithm for $\epsilon = .2$. We report our results in table
\ref{red_t} in the appendix, where we give the ratio on top of the simplified grammar produced
by the clustering algorithm, and the time taken to run just the reduction
rules.

When plotted on a log-log scale, the runtime for our straightforward
implementation of the reduction rules appears to be superpolynomial. While
modest additional gains of about 30 percent are possible for highly
self-similar data, very few additional gains are possible for more diverse
datasets such as gen-n. This appears to be because the only simplifications
that are possible on diverse datasets don't lead to the wide-ranging repetitive
substitutions which allow reduction rules to be applied.

@@

The second piece of text is one of the entries to this year's International
Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c}
. We believe this is a particularly good example since its structure is
foreign, but still discernable. Whitespace is particularly important in
standard english but our lossifiers don't treat it any differently from any
other symbol.  The size of the lossified grammar is 1576 compared with the
original size of 1973. The intricate whitespacing was lost, but the clear
delineation into sections was kept, key words such as `define' and `for' are
very visible, and the texture of the obfuscated C is very obvious. For
instance, after the distinct `\#define' headers, the first section has a bunch
of control statements like `for' and `if', while the middle sections have very
generous spacing and many repeated pointer derence operators such as `***'.
Both of these kinds of textures are retained after lossifying.

@@

@@

ine id, Lt e frargras, e ding
seedn e rut dingrut fes inwhseseed i
in tel, e frt wa

n e frrurhtargras, n dingseedfes inn e dingrut , whseseeds in tel, fe
s inine tht waod3e e frinwee e thir

1e said, e behts ine mamene e 
todividthayfrm e htn e e e frsignd an
seasond anayn ar: 1:n e e e frhts in
e frmamene e  toinhte fr: n t wa

6ine  tcrathts;e grHatoule thay
nde estoule e hte e e e stars al

7ine t e ine mamene e toinht
e fr1:n toule e e ayn e e htd an
todivide htfrm darknes: God tht waod9n e e n frinwee e fr

0e said, binar aundanolye moilngcaledthhath in frwlthay flyabene frentoopen firmamenofmae

1 n God Haedcratwhalesn enry lcalede
